---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load 1,000 rows of data with the ages of people from 18-80

```{r}
set.seed(8675309)

population <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)
```

### Task 3: Sample 100 rows of data from the dataset

```{r}
sample100 <- population |>
  sample_n(100)
```

### Task 4: Calculate the mean age and standard deviation of the sample

```{r}
sample100 |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```

### Task 5: Calculate the mean age and standard deviation of the population

```{r}
population |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```

### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

#The means and standard deviations are very similar to one another: the mean is off by two years, and the standard deviation is off by just over one year. As I calculated below, a smaller sample size also has very similar means and standard deviations, however, they are slightly farther away. That suggests (and I know generally) that the more people that are included as part of a sample, the more accurate the sample will be.

```{r}
sample20 <- population |>
  sample_n(20)
sample20 |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
```

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
