---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

#There is an impressive amount of bloat on this staff, I truly believe that there is no need for there to be this many advisors to the president. The salary values of 0 could potentially cause some problems for this analysis. They are likely to be real outliers. However, I'm sure they will get evened out somewhat by the several salaries close to $200,000.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY, na.rm = TRUE))
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY, na.rm = TRUE), min_salary = min(SALARY, na.rm = TRUE), max_salary = max(SALARY, na.rm = TRUE))

```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

#The fact that the mean and median salaries are very close means the distribution is pretty normal and the spread is relatively even. There aren't any real outliers, but people who are not paid anything are obviously a significant jump away from the lowest full-time salaries given from the White House. The minimum shows that some people are there to give advice for free, the maximum tells me that the size of salaries has not gotten too bloated - just shy of 200K for a major advising role seems reasonable to me. I would predict that there is some sort of payscale they have to stay committed to. I'd also suspect this is some business practice stuff going on where you keep it below the round number for the number to feel more reasonable. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

#My best story idea comes from the sheer volume of advisors in the White House, over 400 feels excessive. I would probably write a story about where their salaries fit in in the percentiles of salaries for all government workers or for people in DC to start to observe this high-volume class of workers who aren't necessarily essential.
