---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
candydata <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```

### Task 3: Sample 5 rows of data from the dataset
https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv

```{r}
sample_5 <- candydata |> sample_n(5)
sample_5
```

### Task 4: Calculate the mean age and standard deviation of the sample, using either the red or black number of M&Ms

```{r}
mean(sample_5$red)
sd(sample_5$red)
```

### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
mean(candydata$red)
sd(candydata$red)
```

### Task 6: Compare the sample and population statistics

#The standard deviations are pretty close to one another, with a difference of 1.2, but the difference in the means is over 8. 

### Task 7: Describe the results

#The standard deviation being 16 means that this data is spread out quite far from its average. The difference in the means being over 8 is significant. That indicates that random sampling is not something that should constantly be relied upon as gospel. These random samples could potentially describe a population in a manner that is very different to reality, especially when there is a large variance within the data as there is here.   

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
